Regression Cube Analysis of Cohort Study Data

### 2015-03-26 draft.tex
start of Sylvias revision
added first part of Sylvias revision

%% indicate IEEE Member or Student Member in form indicated below
\author{Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Katrin Hegenscheid, Henry V{\"o}lzke, Bernard Preim}
\authorfooter{
%% insert punctuation at end of each item
\item
 Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Bernhard Preim are with Otto-von-Guericke University Magdeburg, Germany. E-mail: \{klemm,lawonn,niemann,preim\}@ovgu.de
\item
 Katrin Hegenscheid, Henry V{\"o}lzke are with Ernst-Moritz-Arndt University Greifswald, Germany. E-mail: \{katrin.hegenscheid,voelzke\}@uni-greifswald.de
}

@@

%% Abstract section.
\abstract{%
%\com{Problem}.
Epidemiological studies comprise heterogeneous data about a subject group (a \emph{cohort}) to define disease-specific risk factors.
%%
These data contain information (\emph{features}) about a subjects' lifestyle, medical conditions and also medical image data.
%%
%These features are analyzed using statistical regression analysis to identify feature combinations indicating a disease (the \emph{target feature}).
%%
Statistical regression analysis is used to evaluate these features and to identify feature combinations indicating a disease (the \emph{target feature}).
%%
Although there is a strong demand for overview visualizations of a whole data set towards a target feature, no suitable tool is available for epidemiological researchers.

%\com{New Solution}.
%%
We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
%%
This approach combines all \emph{independent features} with respect to a \emph{target feature} and provides a visualization, which reveals insights into the data by highlighting relationships.
%%
A 3D-visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
%%
Slicing through the \emph{Regression Cube} allows for the detailed analysis of features towards the target disease.
%%
Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
%%
Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
%%
%\com{Validation \& Results}.
%%
We applied our \emph{Regression Cube} method to a hepatic steatosis data set to reproduce results from a data-mining driven analysis.
%%
A qualitative analysis with three domain experts was conducted on a breast fat data set.
%%
We were able to derive new hypotheses about relations between breast fat density and breast lesions towards breast cancer.
%%
%\com{Implications}.
%%
%With our work, we present a visual overview of epidemiological data with the \emph{Regression Cube}, which allows for the first time a interactive regression-based analysis of large feature sets with respect to a disease.
} % end of abstract

@@

\maketitle
%%
Epidemiology aims to characterize health and disease conditions in defined populations (\emph{cohorts}).
%%
Insights about risk factors allow to characterize disease-specific high-risk groups and act as important diagnostic key figures \cite{Fletcher2012}.
%%
They can also be used to give recommendations regarding a healthy lifestyle and provide information about wide-spread diseases.
%%
In the standard workflow, physicians translate observations into hypotheses, which are depicted using epidemiological features and then assessed using regression analyses.

An important epidemiological tool for deriving such features are \emph{cohort studies}, such as the Study of Health in Pomerania (SHIP) \cite{Volzke2011}.
%%
To reduce any selection bias, subjects are invited at random and without a focus on a specific disease.
%%
The acquired features comprise a wide range.
%
Social and lifestyle factors, prior or current diseases and medications as well as medical parameters, such as blood pressure are gathered.
%%
Medical image data from non-radiating modalities. e.g. from magnetic resonance imaging (MRI) is also acquired in modern studies.
%%
These data may be quantified using user-defined landmarks, describing for example shape, volume or diameter of a structure.
%%

Testing features for association with diseases using regression analysis is one of the most important epidemiological tools.
%%
Using it to assess the statistical resilience of a hypotheses rarely involves \emph{more than three features} due to the required subject count.
%%
Due to missing overview techniques, possibly interesting correlations may be missed.
%%
Explorative analyses and overview visualizations of the data set as presented by Klemm et al. \cite{Klemm2014VIS} are not tailored to a specific target variable and mostly highlight correlations between variables, which are known to the domain expert (e.g. correlation between body size and spine shape).
%%
We incorporate the regression analysis, which is familiar to the domain experts into overview visualizations, which can be used for a hypothesis-free analysis or an analysis towards a specific disease or hypotheses.
%%
This is achieved by providing template regression formulas, which are applied to all potential variable combinations.
%%
Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
%%
Difference views between regression formulas allow to assess the influences of individual variables in the process.

Our contributions are:
\begin{itemize}
	\item An overview visualization technique representing feature interactions using target features.
	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
	\item Analysis of confounding variables by providing comparison views between different formula results
	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
\end{itemize}
%%
The term Regression Cube was used before to define for the detailed analysis of one or a small number of regression models \cite{Chan, Ahmadi}.
%%
We refer to it using as visualization technique for comparing and analyzing a large number of different regression models with different input features.
%%
\com{Bernhard, does this suffice as reasoning?}

\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features using statistical methods.
%%
As described by Thew et al. \cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{itemize}
	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of variables depicting the hypothesis and include confounding variables.
	\item Statisticians assess the association of the derived features towards the investigated disease.
\end{itemize}
%%
Relative risks can be determined if a statistical resilient association of features towards a condition are extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS} are used to analyze the data using the classical sequential epidemiological workflow, which uses images largely only to communicate results, rather than providing insight.
%%
An alternative data-driven approach is described as follows.

@@

\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
%\com{Bernhard: Due to overlaps to VIS'14 paper in this section, this can be shortened. E.g. "More discussion of data types and characteristics can be found in [...]"}
%%
\emph{Cohort studies} are a major epidemiological tool for gathering epidemiological features.
%%
They yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either towards a specific disease or with the widest range possible.
%%
The latter allows the data set to be assessed towards different diseases.
%%
The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical image data.
These features are derived through different modalities, such as questionnaires or medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
%\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
\\\\
%\paragraph{Data Types.}
%%
Medical status variables or lifestyle factors are often of \emph{dichotomous} (binary) type.
%%
The data space comprises continuous variables (somatometric variables, such as body weight or BMI or laboratory values) as well as categorical variables (e.g. graduation type).
%%
The data heterogeneity has to be taken into account when the analysis method is chosen.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the variable types and to simplify the method selection.
%%
This is however avoided if possible, since it reduces the information space and also introduces a new information bias, as assumptions are modeled through the discretization.
%%

@@

\paragraph{Confounding Features}
Features influencing the exposure as well as the outcome of an analysis are called confounders.
%%
The analysis model has to be adjusted by normalizing all included features towards the confounding feature.
%%
\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely with increasing age.
%%
It also influences the general body condition and thereby almost all features acquired through cohort studies.
%%
Another important confounder is \emph{Gender}.
%%
Other confounders have to be selected by epidemiologists specific to the investigated condition.

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data.
%%
They are also the foundation of this work.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values describe the resulting function (the \emph{model}).
%%
$R^2$ values describe the goodness of fit; in other words how well the dependent features describe the target feature.
%%
The value is in the range [0,1], where 1 encodes a perfect fit.

@@

\subsection{The Study of Health in Pomerania (SHIP)}
The \texttt{SHIP}, located in Northern Germany aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data per subject.
%%
A second cohort, \texttt{SHIP-TREND} was introduced in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
For the last acquisition (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study ("\emph{method loyality}").
%%
%Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%After the pioneering Rotterdam study (started in 1990), several MR imaging study initiatives were initiated.
%
%They slightly differ in clinical focus, acquired data and epidemiological research questions.
%
%Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%
%Data are collected without focus on a group of diseases.
%
%This allows to query the data regarding many diseases and conditions.
%
%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
Tukey already stated in 1977 that data if too often only analyzed using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to derive hypotheses, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
We put each paper in relation to our work, derive ideas and point out differences, starting with the works closest to ours:
%%
\\\\
The work of Piringer et al. \cite{Piringer}, Chan et al. \cite{Chan}, Angelelli et al. \cite{Angelelli} and Zhang et al. \cite{Zhang2014} is closest to ours regarding different aspects.
%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Akin to our approach, Piringer et al. \cite{Piringer} propose methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to pairwise assess the influence of independent features towards the target.
%%
This is achieved by using a plot matrix displaying models as contours towards the target feature.
%%
They also incorporate 3D visualizations for each pairwise combination, but mainly because it is popular with the engineers in their target domains.
%%
Linked views of model deviations allow to select outlier residuals.
%%
Comparison of models is also included by mapping model values on line-style (e.g. solid for $model_a$ and dashed for $model_b$).
%%
This limits the method to comparing a few models at once.
%%
The plot matrix representation however gets complex and harder to understand with increasing number of independent features.
%%
The main difference is their focus on analyzing one complex model in detail, yielding in extensive plots, while we analyze a large amount of simpler models towards \emph{different} features.
%%
They also focus on metric features, while we process also categorical data.
%%
We share the matrix-arrangement of results depicting relationships between independent features in a model.

@@

\subsection{Case 2: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}
%%
\subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
%%
\textbf{Extracted Hypotheses have to investigated further.}
\com{Tills Feedback - Regression analysis also incorporates assumptions about the data, }
%%

\textbf{Overview Visualizations gain importance with increasing feature count.}
%%
While the classical epidemiological workflow (see Sec.~\ref{sec:Background}) represents the foundation of the research in this area, explorative analysis based on the data gains importance.
%%
Incorporating data mining algorithms with statistically established and overview visualizations allow for reasoning about risk factors using underlying epidemiological data sets.
%%
One major advantage we observed is the confidence of an expert in the results he or she actually observe.
%%
Their participation in the analysis using human pattern-detection and expert knowledge is for experts preferred over results derived using a black-box method.
%%
By observing expected correlations matching the expert knowledge strengthens the confidence in the method and subsequently in the hypothesis generated from unexpected relationships.
%%
%\com{Include the human into the analysis provides more confidence into the result than data mining algorithms. People believe more in what they see, rather than in what they are presented by an algorithm. Regression cubes incorporate this by showing known associations as well as new ones. Preattentive feature display (either through fast slicing through a cube or by showing hot spots) helps.}

>>>>

%% indicate IEEE Member or Student Member in form indicated below
\author{Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Katrin Hegenscheid, Henry V{\"o}lzke, Bernard Preim}
\authorfooter{
%% insert punctuation at end of each item
\item
 Paul Klemm, Kai Lawonn, Sylvia Gla{\ss}er, Uli Niemann, Bernhard Preim are with Otto-von-Guericke University Magdeburg, Germany. E-mail: \{klemm,lawonn,glasser,niemann,preim\}@ovgu.de
\item
 Katrin Hegenscheid, Henry V{\"o}lzke are with Ernst-Moritz-Arndt University Greifswald, Germany. E-mail: \{katrin.hegenscheid,voelzke\}@uni-greifswald.de
}

@@

%% Abstract section.
\abstract{%
%\com{Problem}.
Epidemiological studies comprise heterogeneous data about a subject group (a \emph{cohort}) to define disease-specific risk factors.
%%
These data contain information (\emph{features}) about a subject's lifestyle, medical condition as well as medical image data.
%%
%These features are analyzed using statistical regression analysis to identify feature combinations indicating a disease (the \emph{target feature}).
%%
Statistical regression analysis is used to evaluate these features and to identify feature combinations indicating a disease (the \emph{target feature}).
%%
Although there is a strong demand for overview visualizations of a whole data set towards a target feature, no suitable tool is available for epidemiological researchers.

%\com{New Solution}.
%%
We propose an analysis approach of epidemiological data sets by incorporating all features in an exhaustive regression-based analysis.
%%
This approach combines all \emph{independent features} with respect to a \emph{target feature} and provides a visualization, which reveals insights into the data by highlighting relationships.
%%
A 3D-visualization of all combinations of two to three independent features towards a target acts as an overview of the whole data set, the \emph{Regression Cube}.
%%
Slicing through the \emph{Regression Cube} allows for the detailed analysis of features towards the target disease.
%%
Expert knowledge about disease-specific hypotheses can be included into the analysis by adjusting the regression model formulas.
%%
Furthermore, the influences of features can be assessed using a difference view comparing different calculation results.
%%
%\com{Validation \& Results}.
%%
We applied our \emph{Regression Cube} method to a hepatic steatosis data set to reproduce results from a data-mining driven analysis.
%%
A qualitative analysis with three domain experts was conducted on a breast fat data set.
%%
We were able to derive new hypotheses about relations between breast fat density and breast lesions towards breast cancer.
%%
%\com{Implications}.
%%
%With our work, we present a visual overview of epidemiological data with the \emph{Regression Cube}, which allows for the first time an interactive regression-based analysis of large feature sets with respect to a disease.
} % end of abstract

@@

\maketitle
%%
Epidemiology aims to characterize health and disease conditions in defined populations (\emph{cohorts}).
%%
Insights about risk factors allow to characterize disease-specific high-risk groups and act as important diagnostic key figures \cite{Fletcher2012}.
%%
Furthermore, they can be used to derive recommendations regarding a healthy lifestyle or to provide information about wide-spread diseases.
%%
During the standard workflow, physicians transform observations into hypotheses, which are depicted using epidemiological features and then assessed using regression analyses.

An important epidemiological tool for deriving such features is the evaluation of a \emph{cohort study}, such as the Study of Health in Pomerania (SHIP) \cite{Volzke2011}.
%%
To reduce any selection bias, subjects are randomly invited without a focus on a specific disease.
%%
Hence, a wide range of features is acquired.
%
Social and lifestyle factors, prior or current diseases and medications as well as medical parameters, such as blood pressure, are gathered.
%%
Medical image data from non-radiating modalities, e.g., magnetic resonance imaging (MRI), is also acquired in modern studies.
%%
These data may be quantified using user-defined landmarks, which describe attributes like the shape, the volume or the diameter of a structure.
%%

Testing features for association with diseases using regression analysis is one of the most important epidemiological tools.
%%
Using it to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the required subject count.
%%
% missing overview techniques ist zu hart; es gibt ja techniken (eine scatterplot matrix zb.)
Due to the amount of data and only limited overview techniques, possible correlations may be missed.
%%
Explorative analyses and overview visualizations of the data set as presented by Klemm et al. \cite{Klemm2014VIS} are not tailored to a specific target variable and mostly highlight correlations between variables, which are known to the domain expert (e.g., correlation between body size and spine shape).
%%
We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations, which can be used for a hypothesis-free analysis or an analysis towards a specific disease or hypothesis.
%%
This is achieved by providing template regression formulas, which are applied to all potential variable combinations.
%%
Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
%%
Difference views between regression formulas allow to assess the influences of individual variables on the process.

Our contributions are:
\begin{itemize}
	\item An overview visualization technique representing feature interactions using target features.
	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details on demand techniques for in-depth investigations of feature relationships.
	\item Freely adjustable regression formulas provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
	\item The analysis of confounding variables by providing comparison views between different formula results.
%	\item The open and web-based approach of the system allows for analysis of any data using the presented method.
	\item The expansion and adaption of our method to any data with our open and web-based system.
\end{itemize}
%%
The term Regression Cube has been used for the detailed analysis of one or a small number of regression models \cite{Chan, Ahmadi}.
%%
In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as Regression Cube.
%%
\com{Bernhard, does this suffice as reasoning?}

\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow and requirements.
%%
\subsection{Epidemiological Workflow} \label{EpidemiologicalWorkflow}
Epidemiological research is performed by experts from different academic disciplines, such as physicians, statisticians and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features using statistical methods.
%%
As described by Thew et al. \cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{itemize}
	\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of variables depicting the hypothesis and include confounding variables.
	\item Statisticians assess the association of the derived features towards the investigated disease.
\end{itemize}
%%
Relative risks can be determined if a statistical resilient association of features towards a condition is extracted.
%%
They indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data using the classical sequential epidemiological workflow. Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%
An alternative data-driven approach is described as follows.

@@

\subsection{Epidemiological Data} \label{sec:EpidemiologicalData}
%%
%\com{Bernhard: Due to overlaps to VIS'14 paper in this section, this can be shortened. E.g. "More discussion of data types and characteristics can be found in [...]"}
%%
\emph{Cohort studies} are a major epidemiological tool for gathering epidemiological features.
%%
They yield a highly heterogenous and incomplete information space.
%%
Subject data are collected either towards a specific disease or with the widest range possible.
%%
The latter allows the data set to be assessed towards different diseases.
%%
The feature space comprises information about lifestyle, somatometric variables, medical parameters, genetic data as well as medical image data.
%%
These features are derived through different modalities, such as questionnaires, medical examinations or laboratory analyses.
%%
Many features are sparse, such as follow up questions about a medication or treatment of a certain disease.
%%
Other features are exclusive for a sub-group, such as women-specific questions, e.g., number of born children or period status.
%%
%\com{Restriction of amount of data because it must not be triangulated and because the ethics committees are very restrictive}
\\\\
%\paragraph{Data Types.}
%%
Medical status variables or lifestyle factors are often of \emph{dichotomous} (binary) type.
%%
The data space comprises continuous variables (somatometric variables, such as body weight or BMI or laboratory values) as well as categorical variables (e.g. graduation type).
%%
The data heterogeneity has to be taken into account when the analysis method is chosen.
%%
Continuous data are often discretized (e.g. 10 year steps for age) to equalize the variable types and to simplify the method selection.
%%
However, this should be avoided, since it reduces the information space and also introduces a new information bias, as assumptions are modeled through the discretization.
%%

@@

\paragraph{Confounding Features.}
Features influencing the exposure as well as the outcome of an analysis are called confounders.
%%
The analysis model has to be adjusted by normalizing all included features towards the confounding feature.
%%
\emph{Age} is included as confounder in almost any epidemiological analysis, since most diseases (such as different cancer types) are more likely with increasing age.
%%
It also influences the general body condition and thereby almost all features acquired through cohort studies.
%%
Another important confounder is \emph{Gender}.
%%
Other confounders have to be selected by epidemiologists specific to the investigated condition.

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data.
%%
The presented work is based on regression analysis as well.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values describe the resulting function (the \emph{model}).
%%
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value is in the range [0,1], where 1 encodes a perfect fit.

@@

\subsection{The Study of Health in Pomerania (SHIP)}
The \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%%
It does not focus on a specific disease, making the data set open for many diseases.
%%
Unique for the \texttt{SHIP} is the acquisition of medical image data per cohort subject.
%%
A second cohort, \texttt{SHIP-TREND} was introduced in 2012.
%%
Data for both cohorts are examined in a 5-year time span.
%%
New parameters are added in each iteration, extending the range of investigated diseases.
%%
For the latest two acquisitions (\texttt{SHIP-2} and \texttt{SHIP-TREND-0}), MRI scans are included into the cohort \cite{Hegenscheid2009, Ivanovska2014}.
%%
However, most examinations occur in all stages and are performed according to the same instructions to enable comparisons over the different stages of the study ("\emph{method loyality}").
%%
%Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%After the pioneering Rotterdam study (started in 1990), several MR imaging study initiatives were initiated.
%
%They slightly differ in clinical focus, acquired data and epidemiological research questions.
%
%Starting in 1997 with a cohort of 4,308 subjects, the \texttt{SHIP}, located in Northern Germany, aims to characterize health and disease in the widest range possible \cite{Volzke2011}.
%
%Data are collected without focus on a group of diseases.
%
%This allows to query the data regarding many diseases and conditions.
%
%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
In 1977, Tukey already stated that data is too often only analyzed using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to derive hypotheses, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
We put each paper in relation to our work, derive ideas and point out differences, starting with the works closest to ours:
%%
\\\\
The work of Piringer et al. \cite{Piringer}, Chan et al. \cite{Chan}, Angelelli et al. \cite{Angelelli} and Zhang et al. \cite{Zhang2014} is closest to ours regarding different aspects.
%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Akin to our approach, Piringer et al. \cite{Piringer} propose methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to pairwise assess the influence of independent features towards the target.
%%
This is achieved by using a plot matrix displaying models as contours towards the target feature.
%%
They also incorporate 3D visualizations for each pairwise combination, but mainly because it is popular with the engineers in their target domains.
%%
Linked views of model deviations allow to select outlier residuals.
%%
Comparison of models is also included by mapping model values on line-style (e.g. solid for $model_a$ and dashed for $model_b$).
%%
This limits the method to comparing a few models at once.
%%
The plot matrix representation however gets complex and harder to understand with increasing number of independent features.
%%
The main difference is their focus on analyzing one complex model in detail, yielding in extensive plots, while we analyze a large amount of simpler models towards \emph{different} features.
%%
They also focus on metric features, while we process also categorical data.
%%
We share the matrix-arrangement of results depicting relationships between independent features in a model.

@@

\subsection{Case 2: Hypothesis-driven Analysis of the Hepatic Steatosis Data Set}
%%
\com{1. Can we reconstruct relationships based on the regression cube?}

\com{2. How does the numeric target feature influence the result?}
\subsection{Further Feedback and Lessons Learned} \label{Lessons Learned}
%%
\textbf{Extracted Hypotheses have to investigated further.}
\com{Tills Feedback - Regression analysis also incorporates assumptions about the data, }
%%

\textbf{Overview Visualizations are preferred over Black-box Methods.}
%%
While the classical epidemiological workflow (see Sec.~\ref{sec:Background}) represents the foundation of the research in this area, explorative analysis based on the data gains importance.
%%
Incorporating data mining algorithms with statistically established and overview visualizations allow for reasoning about risk factors using underlying epidemiological data sets.
%%
One major advantage we observed is the confidence of an expert in the results he or she actually observe.
%%
Their participation in the analysis using human pattern-detection and expert knowledge is for experts preferred over results derived using a black-box method.
%%
By observing expected correlations matching the expert knowledge strengthens the confidence in the method and subsequently in the hypothesis generated from unexpected relationships.
%%
%\com{Include the human into the analysis provides more confidence into the result than data mining algorithms. People believe more in what they see, rather than in what they are presented by an algorithm. Regression cubes incorporate this by showing known associations as well as new ones. Preattentive feature display (either through fast slicing through a cube or by showing hot spots) helps.}
